Exercise 2: Quality Attributes and Architectural Design of the HS07 System

### 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
Adding comments to factory diagrams

\section{Quality Attribute Scenarios}
The following chapters documents the selection of the Quality Attributes that we 
select for the HS07 system. The selection of QA is split into a 3-part process.
\begin{itemize}
\item Brainstorming - we think up scenarios
\item Selection - we chose which scenarios are interesting.
\item Refinement - we refine the selected scenarios
\end{itemize}

@@

\begin{table}[!htp]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & Developer updates and tests the code base with a new device type in 1 week. \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Testability  \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} & The developer \\
  \cline{2-3}
  & {\bfseries Stimulus:} & Whishes to add a new device type \\
  \cline{2-3}
  & {\bfseries Artifact} & The code \\
  \cline{2-3}
  & {\bfseries Environment:} & At design time \\
  \cline{2-3}
  & {\bfseries Response:} & Modification is made with no side effects (all unit-/integrationtest pass)\\
  \cline{2-3}
  & {\bfseries Response \mbox{Measure:}} & Modification is made and tested within 1 week \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Does the device use implemented HTTP protocol or do we need an adaptor? Is it necessary to make structual modifications to the user interface? Does the change have architectural impact? \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & Developer is assumed to know the system architecture  \\
  \hline
\end{tabular}
\caption{Refined Scenario \#1 - QA Scenario 9/11/18}
\end{center}
\end{table}

\begin{table}[!htp]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user adds a new actuator and the system configures to use the new hardware automatically within 1 minute \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Usability, Performance \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} & The user \\
  \cline{2-3}
  & {\bfseries Stimulus:} & Attaches a new actuator \\
  \cline{2-3}
  & {\bfseries Artifact} & The system \\
  \cline{2-3}
  & {\bfseries Environment:} & Maintanance operation \\
  \cline{2-3}
  & {\bfseries Response:} & New hardware is automatically added to the system configuration and accessible remotely \\
  \cline{2-3}
  & {\bfseries Response \mbox{Measure:}} & Within one minute \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Is the user allwed to install now components himself? \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
  \hline
\end{tabular}
\caption{Refined Scenario \#2 - QA Scenario 5/21}
\end{center}
\end{table}

\begin{table}[!htp]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user changes the preferred temperature from a web page and the system adapts to it \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Usability,  Availability \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} & The user \\
  \cline{2-3}
  & {\bfseries Stimulus:} & Remotely change the temperature \\
  \cline{2-3}
  & {\bfseries Artifact} & The system \\
  \cline{2-3}
  & {\bfseries Environment:} & Normal operation \\
  \cline{2-3}
  & {\bfseries Response:} & Acknowledge of operation from gateway \\
  \cline{2-3}
  & {\bfseries Response \mbox{Measure:}} & Within 5 seconds 98,9\% of the time, within 30 seconds 1\% of the time and more than 30 seconds 0,1\% of the time (unshceduled down-time) \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & When do the temperature adjustment begin? When is the command is received by the radiator? \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
  \hline
\end{tabular}
\caption{Refined Scenario \#3 - QA Scenario 6/16}
\end{center}
\end{table}

@@

The separation between the gateway and the model packages creates an opportunity for increase of security when that becomes a driving quality attribute, since the gateway information can be hidden within the system, letting the ModelService become the only exposed part.
\begin{figure}[!htb]
\centerline{\epsfig{figure="figures/mv-factory",scale=0.8 }}
\caption{Module View of the abstract factory in the redesigned HS07 system}
\label{fig:mv-factory}
\end{figure}

@@

\begin{figure}[!htb]
\centerline{\epsfig{figure="figures/seq-tfac",scale=0.8 }}
\caption{Sequence Diagram for the Abstract Factory in the redesigned HS07 system}
\label{fig:seq-tfac}
\end{figure}

>>>>

\section{Quality Attribute Scenarios}
The following chapters documents the selection of the Quality Attributes that we 
select for the HS07 system. The selection of QA is split into a 3-part process.
\begin{itemize}
\item Brainstorming - we think up scenarios
\item Selection - we chose which scenarios are interesting.
\item Refinement - we refine the selected scenarios
\end{itemize}
The actual brainstorming and selection (voting) was done in Wiki in order to allow for a decentralised brainstorming and vote. The setup may be seen at:

http://code.google.com/p/pasoos2007hbc/wiki/SAiPM1H2.

All information has been moved to the report so it is only the use of Wiki as a meeting-place that may be of interest.

@@

Table \ref{scenario1}, \ref{scenario2} and \ref{scenario3} shows refined scenarios of the three
quality attribute scenarios that were given the highest priorities in section \ref{sec:selection}.

\begin{table}[!htp]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & Developer updates and tests the code base with a new device type in 1 week. \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Testability  \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} & The developer \\
  \cline{2-3}
  & {\bfseries Stimulus:} & Whishes to add a new device type \\
  \cline{2-3}
  & {\bfseries Artifact} & The code \\
  \cline{2-3}
  & {\bfseries Environment:} & At design time \\
  \cline{2-3}
  & {\bfseries Response:} & Modification is made with no side effects (all unit-/integrationtest pass)\\
  \cline{2-3}
  & {\bfseries Response \mbox{Measure:}} & Modification is made and tested within 1 week \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Does the device use implemented HTTP protocol or do we need an adaptor? Is it necessary to make structual modifications to the user interface? Does the change have architectural impact? \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & Developer is assumed to know the system architecture  \\
  \hline
\end{tabular}
\caption{Refined Scenario \#1 - QA Scenario 9/11/18}
\label{scenario1}
\end{center}
\end{table}

\begin{table}[!htp]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user adds a new actuator and the system configures to use the new hardware automatically within 1 minute \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Usability, Performance \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} & The user \\
  \cline{2-3}
  & {\bfseries Stimulus:} & Attaches a new actuator \\
  \cline{2-3}
  & {\bfseries Artifact} & The system \\
  \cline{2-3}
  & {\bfseries Environment:} & Maintanance operation \\
  \cline{2-3}
  & {\bfseries Response:} & New hardware is automatically added to the system configuration and accessible remotely \\
  \cline{2-3}
  & {\bfseries Response \mbox{Measure:}} & Within one minute \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Is the user allwed to install now components himself? \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
  \hline
\end{tabular}
\caption{Refined Scenario \#2 - QA Scenario 5/21}
\label{scenario2}
\end{center}
\end{table}

\begin{table}[!htp]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user changes the preferred temperature from a web page and the system adapts to it \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Usability,  Availability \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} & The user \\
  \cline{2-3}
  & {\bfseries Stimulus:} & Remotely change the temperature \\
  \cline{2-3}
  & {\bfseries Artifact} & The system \\
  \cline{2-3}
  & {\bfseries Environment:} & Normal operation \\
  \cline{2-3}
  & {\bfseries Response:} & Acknowledge of operation from gateway \\
  \cline{2-3}
  & {\bfseries Response \mbox{Measure:}} & Within 5 seconds 98,9\% of the time, within 30 seconds 1\% of the time and more than 30 seconds 0,1\% of the time (unshceduled down-time) \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & When do the temperature adjustment begin? When is the command is received by the radiator? \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
  \hline
\end{tabular}
\caption{Refined Scenario \#3 - QA Scenario 6/16}
\label{scenario3}
\end{center}
\end{table}

@@

The separation between the gateway and the model packages creates an opportunity for increase of security when that becomes a driving quality attribute, since the gateway information can be hidden within the system, letting the ModelService become the only exposed part.
\clearpage

\begin{figure}[!htb]
\centerline{\epsfig{figure="figures/mv-factory",scale=0.8 }}
\caption{Module View of the abstract factory in the redesigned HS07 system}
\label{fig:mv-factory}
\end{figure}
Figure~\ref{fig:mv-factory} shows the classes involved and their interdependencies for implementing the factory pattern for the sensor module. Please refer to Figure~\ref{fig:seq-tfac} for its runtime usage.

@@

\begin{figure}[!htb]
\centerline{\epsfig{figure="figures/seq-tfac",scale=0.8 }}
\caption{Sequence Diagram for the Abstract Factory in the redesigned HS07 system}
\label{fig:seq-tfac}
\end{figure}
Figure~\ref{fig:seq-tfac} shows initialization sequence when using the SensorFactory illustrated in Figure~\ref{fig:mv-factory}. The sequence shows the use of the factory to create a decoupling between the actual Thermometer implementation and the ThermometerService.
\clearpage
